Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{DISTRIBUTION STATEMENT A. Approved for public release.}

### 2022-09-19 paper/abstract.tex
done with abstract and intro now

Cyber-physical systems, such as avionics, must be tolerant to
cyber-attacks in the same way they are tolerant to random faults: they
must gracefully recover, or safely shut down, as requirements dictate.
We have developed a workflow for creating, and inserting,
high-assurance components implementing cyber-resiliency into a
model-based systems engineering environment.  Example high-assurance
components are filters, which guard against malformed input, and
runtime monitors, which guard against spoofing and other malicious
behavior. A formal specification in the form of a \emph{code contract}
defines each high-assurance component and is developed with the
support of \emph{test contracts} for testing the specified behavior.
Once tested, model checking used to verify that the added
high-assurance component indeed addresses system-level cyber
requirements.  Implementations for these high-assurance components are
directly synthesized from their code contracts and are backed up by
proofs showing that high-level specifications map in a
semantics-preserving way to code generated by a verified compiler.  We
report on a case study that cyber-hardened a UAV system by inserting
high assurance components to harden the open source Air Force Research
Laboratory's OpenUxAS services for route planning.  The case study
demonstrates that synthesizing correct implementations from code
contracts is feasible in real-world systems engineering.

>>>>

%% Cyber-physical systems, such as avionics, must be tolerant to
%% cyber-attacks in the same way they are tolerant to random faults: they
%% must gracefully recover, or safely shut down, as requirements dictate.

We have developed a workflow for creating and inserting high-assurance
components that enforce cyber-resiliency requirements into a
model-based systems engineering environment.  Example high-assurance
components are filters, which guard against malformed input, and
runtime monitors, which guard against spoofing and other malicious
behavior. A formal specification in the form of a \emph{code contract}
defines each high-assurance component and is developed with the
support of \emph{test contracts} for testing the specified behavior.
Model checking is used to verify that an inserted high-assurance
component indeed addresses system-level cyber
requirements. Implementations for high-assurance components are
directly synthesized from their code contracts and are backed up by
proofs showing that high-level specifications map in a
semantics-preserving way to code generated by a verified compiler.  We
report on a case study that cyber-hardened a UAV system by inserting
high assurance components to harden the open source Air Force Research
Laboratory's OpenUxAS services for route planning.  The case study
demonstrates that synthesizing correct implementations from code
contracts is feasible in real-world systems engineering.

### 2022-09-19 paper/intro.tex
done with abstract and intro now

\egm{Add to introduction and related work, that what is done at
Galois called Copilot. Its a stream processing langague that generates
code from the specification. Copilot to C. Effectively CodeGen from
Lustre. Bounds how far needed to look in past for any value.}

@@

We have been developing the {\em BriefCASE} toolsuite to address this need.
\brfcs\ is a Model-Based Systems Engineering (MBSE) environment
built in the Open Source AADL\footnote{AADL is the acronym for
Architecture Analysis and Design Language~\cite{aadl}.}  Tool
Environment (OSATE) to add new design, analysis, and code generation
capabilities for building cyber-resilient systems.

In this paper we describe how \brfcs\ facilitates inserting,
specifying, testing, and synthesizing high assurance components into a
system to improve its cyber-resiliency.  The main organizing concept
is that of an architecture-to-architecture \emph{security-improving
transform}, achieved via the insertion of a new architectural
component aimed at mitigating a cyber-vulnerability.  We describe two
cyber-resiliency transformations in this paper: (1) the insertion of a
filter to prevent malformed data from a malicious actor being
propagated to downstream components, and (2) the insertion of a
runtime monitor to detect (and alert) unexpected behaviors arising
from untrusted components.

A code contract is a formal specification in the Assume Guarantee
Reasoning Environment (\agr) language.
\agr\ is a compositional reasoning verification engine that uses \emph{contracts} on components to specify input and output properties and then prove whether or not those properties hold when given a sub-component implementation \cite{agree2013}.
The code contract language is Turing complete allowing the designer to
specify arbitrarily complex behavior.  These contracts are unit tested
for correctness with \emph{test contracts}.  Test contracts define
test scenarios to be implemented by the code contract under test.
\agr\ proves whether or not that code contract implementation is correct, enabling the designer to iteratively test the component behavior inside the \brfcs\ environment.
Once the behavior of the code contract is verified, \agr\ proves that---due to the newly
included high-assurance components---the hardened system meets its
cyber-resiliency requirements.

Another novel aspect of our workthe approach discussed in this manuscript is
the synthesis of the code contracts for the high-assurance components
to \ckml, a verified compiler implementation for the functional
programming language ML \cite{cakeml}.  This manuscript describes in
detail the synthesis path from code contracts to \ckml\ code,
providing a formal framework in which to argue correctness. \ckml\
then provides a verified compilation path to several different target
binaries (and also proving that the meaning of the \ckml\ source code
is exactly preserved in the final binaries).  A key contribution here
is that the code contract semantics are defined in such a way that
the \agr\ verification results for the code contract hold for the
deployed component, \ie, the component will detect and prevent the
indicated cyber-vulnerabilities over all possible finite
inputs. Preliminary work has shown how to lift this result to infinite
input traces as these systems are inherently reactive and intended to
run forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

The manuscript further details a case study applying these
transformations with \brfcs\ to an Unmanned Aerial Vehicle (UAV)
system that uses the Air Force Research Laboratory's OpenUxAS services
for route planning.  OpenUxAS, as an open source product, is
considered \emph{untrusted}.  The UAV system is thus transformed to be
resilient to malicious behavior that may arise in the untrusted
component.  Here the transforms add filters to guard against malformed
input and monitors to guard against malicious flight plans from
OpenUxAS. The case study system is complex and shows the viability of
the approach in potential full-scale industrial design.

\brfcs\ is open source and publicly available \cite{fmide} as are the examples and case study discussed in this manuscript \cite{repo, phase2, camkes, case}.
Our approach currently applies to common cyber-vulnerabilities, such
as overflow, lack of input validation, and supply-chain issues;
however, other cyber-vulnerabilities such as side-channel attacks and
denial of service are not yet dealt with in our work.  Here, we do not
report on the invention of a new type of high-assurance component in
terms of capability; instead, our contribution is in the automated
synthesis of security-improving components from formal specifications
and a means to show that the synthesis is correct.

We now give a summary of the contributions detailed in this
manuscript.
\begin{compactitem}
  \item The language and semantics of code contracts to specify the behavior of high assurance components.
  \item Test contracts as a stylized mechanism to unit test code contracts inside the \brfcs\ framework with \agr.
  \item A synthesis path from code contracts to \ckml\ with a formal framework to argue correctness.
  \item A formal argument that \agr\ verification of code contracts carry over to the resulting binaries from the \ckml\ compiler.
  \item A case study that used the implementation of the approach in \brfcs\ to add cyber-resiliency to a non-trivial UAV system.
\end{compactitem}
The rest of this manuscript is organized as follows.
\secref{sec:overview} is an overview of the \brfcs\ environment and key tools in that environment relative to the contributions of this work. The approach is illustrated by a simple example in
\secref{sec:example}. Essential background on \agr\ verification and specification are
presented in \secref{sec:agree}.
The language and semantics of code contracts are defined in \secref{sec:code-contracts}, and test contracts are defined in \secref{sec:testing}.
The synthesis pathway is covered in Section~\ref{sec:synthesis}.
\secref{sec:case-study} discusses the case study.
This is followed by related work in \secref{sec:related-work}.
The conclusions and future work are in \secref{sec:conclusion}.

>>>>

%% \egm{Add to introduction and related work, that what is done at
%% Galois called Copilot. Its a stream processing langague that generates
%% code from the specification. Copilot to C. Effectively CodeGen from
%% Lustre. Bounds how far needed to look in past for any value.}

@@

We have been developing the {\em BriefCASE} toolsuite to address this need.
\brfcs\ is a Model-Based Systems Engineering (MBSE) environment
built in the Open Source AADL\footnote{AADL is the acronym for
Architecture Analysis and Design Language~\cite{aadl}.}  Tool
Environment (OSATE) to add new capabilities for building
cyber-resilient systems.

In this paper we describe how \brfcs\ facilitates specifying, testing,
verifying, and synthesizing high assurance components inserted into a
system to improve its cyber-resiliency.  The main organizing concept
is that of an architecture-to-architecture \emph{security-improving
transform}, achieved via the insertion of a new architectural
component aimed at mitigating a cyber-vulnerability.  We describe two
such transformations in this paper: (1) the insertion of a
filter to prevent malformed data from a malicious actor being
propagated to downstream components, and (2) the insertion of a
runtime monitor to detect (and alert) unexpected behaviors arising
from untrusted components.

{\agr}
\footnote{AGREE is the acronym for \emph{Assume Guarantee Reasoning Environment} \cite{agree2013}.}
provides the formal setting for our work.
{\agr} is implemented in OSATE as an \emph{annex}; hence AGREE
specifications inherit AADL constructs for describing system
structure, interconnections, and data. {\agr} uses \emph{contracts}
expressed in a variant of Past-Time Linear Temporal Logic to formally
specify input and output properties of components and systems. In this
paper we introduce the notion of a \emph{code contract}: a class of
contracts that support the synthesis of implementations. The code
contract language is Turing complete, allowing the designer to specify
arbitrarily complex behavior.  These contracts are unit tested for
correctness with \emph{test contracts}.  Test contracts define test
scenarios to be implemented by the code contract under test.  {\agr}
proves whether or not that code contract implementation is correct,
enabling the designer to iteratively test the component behavior
inside the \brfcs\ environment.  Once the behavior of the code
contract is verified, \agr\ proves that---due to the newly included
high-assurance components---the hardened system meets its
cyber-resiliency requirements.

A novel aspect of our work is the synthesis of {\ckml} code from the
code contracts for high-assurance components. {\ckml} is a dialect of
the ML programming language possessing a verified
compiler \cite{cakeml}.  We give a detailed outline of the synthesis
path from code contracts to \ckml\ code, providing a formal framework
in which to argue correctness. \ckml\ then provides a verified
compilation path to several different target binaries (and also
proving that the meaning of the \ckml\ source code is exactly
preserved in the final binaries).  A key contribution here is that the
code contract semantics are defined in such a way that the \agr\
verification results for the code contract hold for the deployed
component, \ie, the component will detect and prevent the indicated
cyber-vulnerabilities over all possible finite inputs. Preliminary
work has shown how to extend our work to handle infinite input traces
as these systems are inherently reactive and intended to run
forever~\cite{case-verified-filter}, \cite{cakeml-space-cost}.

We report on a case study applying these transformations with \brfcs\
to an Unmanned Aerial Vehicle (UAV) system that uses the Air Force
Research Laboratory's OpenUxAS services for route planning.  OpenUxAS,
as an open source product, is considered \emph{untrusted}.  The UAV
system is thus transformed to be resilient to malicious behavior that
may arise in the untrusted component.  In particular, filters are
added to guard against malformed input and a monitor is created to
guard against malicious flight plans generated from OpenUxAS. The case
study system is complex and shows the viability of the approach in
potential full-scale industrial design.

{\brfcs} is open source and publicly available \cite{fmide}, as are the
examples and case study we discussed in this paper \cite{repo,
phase2, camkes, case}.

\subsubsection*{Contributions}
Our approach currently applies to common cyber-vulnerabilities, such
as overflow, lack of input validation, and supply-chain issues;
however, other cyber-vulnerabilities such as side-channel attacks and
denial of service are not yet dealt with in our work.  Here, we do not
report on the invention of a new type of high-assurance component in
terms of capability; instead, our contribution is in the automated
synthesis of security-improving components from formal specifications
and a means to show that the synthesis is correct. We see the
following as the contributions of this paper:

\begin{compactitem}
  \item The language and semantics of code contracts to specify the behavior of high assurance components.
  \item Test contracts as a stylized mechanism to unit test code contracts inside the \brfcs\ framework with \agr.
  \item A synthesis path from code contracts to \ckml\ with a formal framework to argue correctness.
  \item A formal argument that \agr\ verification of code contracts carry over to the resulting binaries from the \ckml\ compiler.
  \item A case study that uses \brfcs\ to add cyber-resiliency to a non-trivial UAV system.
\end{compactitem}

The rest of the paper is organized as follows.
\secref{sec:overview} is an overview of the {\brfcs} environment.
Our approach is illustrated by a simple example in
\secref{sec:example}. Essential background on {\agr} verification and specification are
presented in \secref{sec:agree}.  The language and semantics of code
contracts are defined in \secref{sec:code-contracts}, and test
contracts are described in \secref{sec:testing}.  The synthesis pathway
is covered in Section~\ref{sec:synthesis}.
\secref{sec:case-study} discusses the case study.
This is followed by related work in \secref{sec:related-work}.
The conclusions and future work are in \secref{sec:conclusion}.
